---
title: 'Online Resource 2'
author: "Rachael E. Bonoan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R code for statistical analysis of phenology and population metrics for __Changes in phenology and abundance of an at-risk butterfly__.

__Authors:__ Rachael E. Bonoan, Elizabeth Crone, Collin B. Edwards, Cheryl B. Schultz

## Load packages
```{r, warning=FALSE, message=FALSE}
library(plyr)
library(tidyverse)
library(MASS)
library(car)
library(emmeans)
library(multcompView)
library(here)
library(ggplot2)
library(patchwork)

library(broom)
library(reshape2)
library(ggcorrplot)
```

## Themes and color palette for graphs
```{r}
# color palette
pal_9sites_mono2=c("#7570b3", 
                    "#8d3e01", "#c05402",  "#f26a02", "#fd862a", 
                    "#105d46", "#178867",  "#1fb487", "#2bdba6")

# theme for trends over time
pheno.theme =  theme(panel.border = element_rect(colour = NA,fill=NA),
                       panel.background=element_blank(),
                       panel.grid.major=element_blank(),
                       panel.grid.minor=element_blank(),
                       axis.line=element_line(colour="black"),
                       axis.text.y=element_text(colour="black",size=13),
                       axis.text.x=element_text(colour="black",size=13, angle=45, vjust=0.5),
                       axis.ticks=element_line(colour="black"),
                       axis.title.y=element_text(colour="black",size=14),
                       axis.title.x=element_text(colour="black",size=14),
                       legend.text=element_text(size=10),
                       legend.title=element_text(size=12),
                       aspect.ratio = 1)

# theme for association between pop growth and phenology
slope.theme =  theme(panel.border = element_rect(colour = NA,fill=NA),
                     panel.background=element_blank(),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     axis.line=element_line(colour="black"),
                     axis.text.y=element_text(colour="black",size=13),
                     axis.text.x=element_text(colour="black",size=13),
                     axis.ticks=element_line(colour="black"),
                     axis.title.y=element_text(colour="black",size=14),
                     axis.title.x=element_text(colour="black",size=14),
                     #legend.position = "bottom",
                     legend.text=element_text(size=10),
                     legend.title=element_text(size=12),
                     aspect.ratio = 1)

```

# TRENDS OVER TIME (years)

## Data: estimates from GLMM curves

### Import
ALL Variables (__bold__ = variables of interest for THIS analysis)

* __scenario:__ name of the fitting. Site name, followed by "base" for using only the raw data, and "monthly" for the cases where we added observations of 0 counts on days 15, 46, 74, 227, 258, 288, 319, 349 of each year.
* __year:__ year of fitting
* b0, b1, b2: fitted coefficients, after unscaling. Corresponds to the linear model presented in the methods paper.
* __mu:__ estimated mean day of activity
* __fp:__ estimated 80% flight period (80FP; days from 0.1 quantile and 0.9 quantile of activity)
* __N:__ estimated abundance index 
* __tenperc:__ estimated "start of activity" (day of 0.1 quantile of activity)
* h: estimated peak activity. This is not a metric we have focused on, but I have included because I had the code.
* __ofs:__ estimated observed flight season (OFS; days between first and last estimate of N = 1)
* badfit: Were the best-fitting estimates biologically implausible? (exponential decay or growth instead of gaussian)
```{r, warning=FALSE, message=FALSE}
# data downloaded from Dropbox, 4_res (file has same name)
dat = read_csv(here::here("4_res","yearly-summary-full.csv"))
head(dat)
```

### Process
Two variables processed before analysis:

__scenario:__ divided into three variables

* cons.area: conservation area (Fir = Fir Butte, WC = Willow Creek, FR = Fern Ridge)
* name: name of site (9 sites), filtered to remove Hayfield 
* scenario: see above, filtered for only the base scenario

__N:__ two mutations to calculate __Nln__

1. N, N.025, N.975 all divided by 10 (FBB lifespan) to convert "activity days" to butterfly population estimate

2. AFTER dividing by 10, all "Ns" natural-logged

_Data also reorderd for consistency with graph legends and comparions._

```{r, warning=FALSE, message=FALSE}
dat2 = dat %>% 
  filter(site != "Hayfield") %>%
  mutate(bfly = N / 10) %>%
  mutate(Nln = log(bfly)) # HAVE TO DIVDE BY 10 FIRST!


head(dat2)
#summary(dat2)
dat2$name = gsub(".*_", "", dat2$site)
dat2$cons.area = gsub("_.*", "", dat2$site)
# reorder for consistency (alphabetical, by conservation area)
dat2$name = factor(dat2$name, levels = c ("Butte","Eaton", "GreenOaks", "Shore", "Spires","Bailey", "FirGrove", "Main", "North"))
```

## Data analysis 

### Population size (Nln) over time (years)
Fixed effects: year, name, interaction
```{r}
mpop = lm(Nln ~ as.numeric(year) * name, data=dat2)
hist(resid(mpop)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mpop) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
 summary(mpop)
```

#### emmeans() and CLD()

Compare all sites, no correction for multiple comparisons, interaction effect
```{r, message=FALSE, warning=FALSE}
# significant interaction, report for graph
em.mpop = emtrends(mpop, pairwise ~ name, var = "year", adjust = "none")
#em.mpop

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(em.mpop, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "none")

```

Compare all sites, Tukey's correction for multiple comparisons, interaction effect (slopes)
```{r, message=FALSE, warning=FALSE}
# significant interaction, report for graph
tuk.mpop = emtrends(mpop, pairwise ~ name, var = "year", adjust = "Tukey")
#tuk.mpop

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(tuk.mpop, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "Tukey")
```

#### Graph (Figure 3, panel a)
```{r, message=FALSE}
pop=ggplot(dat2, aes(x=year, y=Nln, shape = cons.area, color = name))+
    geom_point(size=2, alpha = 0.4)+
    geom_smooth(method=lm,se=FALSE,size=0.7)+
    #geom_linerange(aes(ymin = N.025ln, ymax = N.975ln)) + 
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed")+
    stat_ellipse(aes(x = year, y = Nln, group = NA), type = "norm", color="black", linetype="dashed", size=1)+
    #facet_grid(~cons.area)+
    xlab("Year")+
    ylab("ln(Estimated population size)")+
    labs(color="Site")+
    ylim(0, 10) +
    pheno.theme+
    scale_x_continuous(breaks=c(1990, 1995, 2000, 2005, 2010, 2015, 2020))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_color_manual(values=pal_9sites_mono2,
                       labels=c("Fir Butte", "Eaton, FR", "Green Oaks, FR", "Shore, FR",
                                "Spires, FR", "Bailey, WC", "Fir Grove, WC", "Main, WC", "North, WC"))+
    annotate("text", x = 2016, y = 2, label = "r = 0.70")
pop
```

### Day of peak flight (DOY) over time (years)
Fixed effects: year, name, interaction
```{r}
mmu = lm(mu ~ year * name, data=dat2)
hist(resid(mmu)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mmu) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
 summary(mmu)
```

#### mean change per year
```{r}
emtrends(mmu, "year", var = "year")
```

#### emmeans() and CLD()
compare all sites, no correction for multiple comparisons
```{r, message=FALSE, warning=FALSE}
# no significant interaction, use emmeans()
em.mmu = emmeans(mmu, specs = pairwise ~ name, adjust = "none")
em.mmu
#em.mmu

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(em.mmu, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "none")
```

Compare all sites, Tukey's correction for multiple comparisons, site effect (means)
_Will not look at multiple comparisons for slopes because no significant interaction effect_
```{r, message=FALSE, warning=FALSE}
tuk.mmu = emmeans(mmu, specs = pairwise ~ name, adjust = "Tukey")
#tuk.mmu

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(tuk.mmu, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "Tukey")
```

#### Graph (Figure 3, panel b)
```{r, message=FALSE}
peak=ggplot(dat2, aes(x=year, y=mu, shape = cons.area, color = name))+
    geom_point(size=2,alpha=0.4)+
    geom_smooth(method=lm,se=FALSE,size=0.7)+
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed")+
    stat_ellipse(aes(x = year, y = mu, group = NA), type = "norm", color="black", linetype="dashed", size=1)+
    #facet_grid(~cons.area)+
    xlab("Year")+
    ylab("Day of peak flight (DOY)")+
    labs(color="Site")+
    pheno.theme+
    scale_x_continuous(breaks=c(1990, 1995, 2000, 2005, 2010, 2015, 2020))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_color_manual(values=pal_9sites_mono2,
                       labels=c("Fir Butte", "Eaton, FR", "Green Oaks, FR", "Shore, FR",
                                "Spires, FR", "Bailey, WC", "Fir Grove, WC", "Main, WC", "North, WC"))+
    annotate("text", x = 2020, y = 123, label = "r = 0.27")
  
peak
```

### 80% flight period (80FP) over time (years)
Fixed effects: year, name, interaction
```{r}
mfp = lm(fp ~ year * name, data=dat2)
hist(resid(mfp)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mfp) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
 summary(mfp)
```

#### emmeans() and CLD()
compare all sites, no correction for multiple comparisons
```{r, message=FALSE, warning=FALSE}
# no significant interaction, use emmeans()
em.mfp = emmeans(mfp, pairwise ~ name, adjust = "none")
#em.mfp

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(em.mfp, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "none")
```

Compare all sites, Tukey's correction for multiple comparisons
```{r, message=FALSE, warning=FALSE}
# no significant interaction, use emmeans()
tuk.mfp = emmeans(mfp, pairwise ~ name, adjust = "Tukey")
#tuk.mfp

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(tuk.mfp, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "Tukey")
```

#### Graph (Figure 3, panel c)
```{r}
period=ggplot(dat2, aes(x=year, y=fp, shape = cons.area, color = name))+
    geom_point(size=2,alpha=0.4)+
    geom_smooth(method=lm,se=FALSE,size=0.7)+
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed")+
    stat_ellipse(aes(x = year, y = fp, group = NA), type = "norm", color="gray", linetype="dashed", size=1)+
    #facet_grid(~cons.area)+
    xlab("Year")+
    ylab("Length of 80FP (days)")+
    labs(color="Site")+
    pheno.theme+
    scale_x_continuous(breaks=c(1990, 1995, 2000, 2005, 2010, 2015, 2020))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_color_manual(values=pal_9sites_mono2,
                       labels=c("Fir Butte", "Eaton, FR", "Green Oaks, FR", "Shore, FR",
                                "Spires, FR", "Bailey, WC", "Fir Grove, WC", "Main, WC", "North, WC"))+
  annotate("text", x = 2017, y = 9.5, label = "r = 0.07")

period
```

### Observed flight season (OFS) over time (years)
Fixed effects: year, name, interaction
```{r}
mofs = lm(ofs ~ year * name, data=dat2)
hist(resid(mofs)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mofs) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
 summary(mofs)
```

#### emmeans() and CLD()
compare all sites, no correction for multiple comparisons
```{r, message=FALSE, warning=FALSE}
# significant interaction, use emtrends()
em.mofs = emtrends(mofs, pairwise ~ name, var = "year", adjust = "none")
#em.ofs

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(em.mofs, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "none")
```

Compare all sites, Tukey's correction for multiple comparisons of interaction effect (slope)
```{r, message=FALSE, warning=FALSE}
# significant interaction, use emtrends()
tuk.mofs = emtrends(mofs, pairwise ~ name, var = "year", adjust = "Tukey")
tuk.mofs

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(tuk.mofs, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "Tukey")
```

Compare all sites, Tukey's correction for multiple comparisons of site effect (means)
```{r, message=FALSE, warning=FALSE}
# significant interaction, use emtrends()
tuk.mofs.mean = emmeans(mofs, pairwise ~ name, adjust = "Tukey")
#tuk.mofs.mean

# assign letters for sig groups
## Note: CLD is now deprecated. 
# CLD(tuk.mofs.mean, details = TRUE, sort = FALSE, alpha = 0.05, Letters = c("a", "b", "c", "d", "e", "f"), adjust = "Tukey")
```

#### Graph (Figure 3, panel d)
```{r, warning=FALSE}
season=ggplot(dat2, aes(x=year, y=ofs, shape = cons.area, color = name))+
    geom_point(size=2,alpha=0.4)+
    geom_smooth(method=lm,se=FALSE,size=0.7)+
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed")+
    stat_ellipse(aes(x = year, y = ofs, group = NA), type = "norm", color="black", linetype="dashed", size=1)+
    #facet_grid(~cons.area)+
    xlab("Year")+
    ylab("Length of OFS (days)")+
    labs(color="Site")+
    pheno.theme+
    scale_x_continuous(breaks=c(1990, 1995, 2000, 2005, 2010, 2015, 2020))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_color_manual(values=pal_9sites_mono2,
                       labels=c("Fir Butte", "Eaton, FR", "Green Oaks, FR", "Shore, FR",
                                "Spires, FR", "Bailey, WC", "Fir Grove, WC", "Main, WC", "North, WC"))+
    annotate("text", x = 2016, y = 17.5, label = "r = 0.43")

season
```

### Put graphs together for FIGURE 3
A bit hard to see all panels in the knitted document, refer to graphs above. Letters denoting significance added in Adobe Illustrator. "r" aslo changed to "R^2" in Illustrator.
```{r, eval=FALSE}
(pop + peak + period + season) +
    plot_layout(guides="collect", ncol = 2) +
    plot_annotation(tag_levels = 'A') &
    theme(plot.tag = element_text(size = 14))
```

# ASSOCIATION BETWEEN PHENO AND POP GROWTH

## Data: slopes of estimates over time (years)

### Import
ALL Variables (__bold__ = variables of interest for THIS analysis)

* __scenario:__ name of the fitting. Site name, followed by "base" for all cases (code is designed to be flexible enough for multiple analyses approaches).

* __vals.mu:__ day of mean activity for this site, averaged across all years

* __vals.fp:__ activity period (days between 0.1 and 0.9 quantile of activity), averaged across all years

* __vals.N:__: abundance index average across all years

* __slope.mu, slope.mu.025, slope.mu.975:__ change in day of activity through time (linear regression of mu vs year)

* __slope.fp5:__ change in 80FP through time (linear regression of fp vs year)

* __slope.ofs:__ change in OFS through time (linear regression of ofs vs year)

* __slope.logN:__ change in log abundance index through time (linear regression of log(N) vs time).


```{r, warning=FALSE, message=FALSE}
# data downloaded from Dropbox, 4_res (file has same name)
dat.cor = read_csv(here::here("4_res","sitely-summary-full.csv"))
head(dat.cor)
```

### Process
On variable processed before analysis.

__scenario:__ divided into two variables

* cons.area: conservation area (Fir = Fir Butte, WC = Willow Creek, FR = Fern Ridge)
* name: name of site (9 sites)

_Data also reorderd for consistency with graph legends and comparions._
```{r, warning=FALSE, message=FALSE}
dat.cor2 = dat.cor %>% 
  separate(scenario, c("cons.area", "name"))
head(dat.cor2)
#summary(dat.cor2)

dat.cor2$name = factor(dat.cor2$name, levels = c ("Butte", 
                                          "Eaton", "GreenOaks", "Shore", "Spires",
                                          "Bailey", "FirGrove", "Main", "North"))

```

## Correlation plots
Quick look to see general relationships, will run linear models (below) for relationships we're interested in interpretting. 
```{r, eval=FALSE}
dat.cor2.test = dat.cor2 %>%
  dplyr::select(vals.mu, vals.fp, vals.N, vals.ofs,
         slope.mu, slope.fp, slope.logN, slope.ofs)

dat.cor2.test = dat.cor2 %>%
  dplyr::select(-cons.area, -name)


panel.hist <- function(x, ...){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE, breaks=20)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
panel.cor <- function(x, y,
                      digits = 2,
                      prefix = "",
                      cex.cor,
                      ...){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y, use="complete.obs"))
  rsigned=cor(x, y, use="complete.obs")
  p = cor.test(x, y, use="complate.obs")$p.value
  txt <- format(c(rsigned, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  col="grey"
  if(p<.1){col="black"}
  if(p<.05){col="blue"}
  text(0.5, 0.5, txt, cex = cex.cor *(r+.2)/(r+1), col=col)
}
panel.cortest = function(x,y,
                         digits=2,
                         prefix="",
                         cex.cor,
                         ...){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  p <- signif(cor.test(x, y, use="complete.obs")$p.value,2)
  if(missing(cex.cor)) cex.cor <- 2
  col="grey"
  if(p<.1){col="black"}
  if(p<.05){col="red"}
  text(.5, .5, p, cex=cex.cor, col=col)
}

pairs(dat.cor2.test,
      lower.panel=panel.cor,
      diag.panel=panel.hist)
```

## Change in Nln vs. change in mu
Fixed effects: change in peak DOY (mu)
```{r}
mchange.mu = lm(slope.logN ~ slope.mu, data=dat.cor2)
hist(resid(mchange.mu)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mchange.mu) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
summary(mchange.mu)
```

#### Graph (Figure 4, panel a)
```{r, warning=FALSE}
change.mu=ggplot(dat.cor2, aes(x=slope.mu, y=slope.logN, color = cons.area, shape = cons.area))+
    geom_point(size = 3.5)+
    #geom_errorbarh(aes(xmin = slope.mu.025, xmax = slope.mu.975), height = 0, size = 0.5)+
    #geom_linerange(aes(ymin = slope.logN.025, ymax = slope.logN.975), width = 0, size = 0.5)+
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed") + 
    stat_ellipse(aes(x = slope.mu, y = slope.logN, group = NA), type = "norm", color="gray", linetype="dashed", size=1)+
    #facet_wrap(~cons.area)+
    xlab("Change in day of peak flight")+
    ylab("Population growth rate")+
    ylim(-0.25, 0.5)+
    xlim(-0.8, 0)+
    slope.theme+
    #geom_text_repel(aes(color=factor(cons.area), label=name))+
    scale_color_manual(name = "Conservation area",
                       values=c("#7570b3", "#d95f02", "#1b9e77"),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    annotate("text", x = -0.05, y = 0.4, label = "r = 0.001")+
    annotate("rect", xmin = -0.8, xmax = 0, ymin = -0.25, ymax = 0, alpha = 0.1) 
    # + geom_text_repel(aes(label=name), color = "black")
change.mu
```

## Change in Nln vs. change in 80FP
Fixed effects: change in 80% flight period (80FP)
```{r}
mchange.fp = lm(slope.logN ~ slope.fp, data=dat.cor2)
hist(resid(mchange.fp)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mchange.fp) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
summary(mchange.fp)
```

#### Graph (Figure 4, panel b)
```{r, warning=FALSE}
change.fp=ggplot(dat.cor2, aes(x=slope.fp, y=slope.logN, color=cons.area, shape = cons.area))+
    geom_point(size=3.5)+
    #geom_errorbarh(aes(xmin = slope.fp.025, xmax = slope.fp.975), height = 0, size = 0.5) +
    #geom_errorbar(aes(ymin = slope.logN.025, ymax = slope.logN.975), width = 0, size = 0.5)+
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed") + 
    stat_ellipse(aes(x = slope.fp, y = slope.logN, group = NA), type = "norm", color="gray", linetype="dashed", size=1)+
    #facet_wrap(~cons.area)+
    xlab("Change in length 80FP")+
    ylab("Population growth rate")+
    labs(color="Conservation Area")+
    ylim(-0.25, 0.5)+
    xlim(-0.2,0.3)+
    slope.theme+
    #geom_text_repel(aes(color=factor(cons.area), label=name))+
    scale_color_manual(name = "Conservation area",
                       values=c("#7570b3", "#d95f02", "#1b9e77"),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    annotate("text", x = 0.25, y = 0.4, label = "r = 0.05")+
    annotate("rect", xmin =-0.2, xmax = 0.3, ymin = -0.25, ymax = 0, alpha = 0.1)
    # + geom_text_repel(aes(label=name), color = "black") 
change.fp
```

## Change in Nln vs. change in OFS
Fixed effects: change in observed flight season (OFS)
```{r}
mchange.ofs = lm(slope.logN ~ slope.ofs, data=dat.cor2)
hist(resid(mchange.ofs)) # check residuals for model
```

#### Anova()
gives signficance values for fixed effects
```{r}
Anova(mchange.ofs) 
```

#### summary()
gives adjusted R-square for full linear regression
```{r}
summary(mchange.ofs)
```

#### Graph (Figure 4, panel c)
```{r, warning=FALSE}
change.ofs=ggplot(dat.cor2, aes(x=slope.ofs, y=slope.logN, color=cons.area, shape = cons.area))+
    geom_point(size=3.5)+
    #geom_errorbarh(aes(xmin = slope.ofs.025, xmax = slope.ofs.975), height = 0, size = 0.5) +
    #geom_errorbar(aes(ymin = slope.logN.025, ymax = slope.logN.975), width = 0, size = 0.5)+
    #stat_smooth(method=lm,se=TRUE,color="black",linetype="dashed") + 
    stat_ellipse(aes(x = slope.ofs, y = slope.logN, group = NA), type = "norm", color="black", linetype="dashed", size=1)+
    #facet_wrap(~cons.area)+
    xlab("Change in length OFS")+
    ylab("Population growth rate")+
    labs(color="Conservation Area")+
    ylim(-0.25, 0.5)+
    xlim(-1.25, 2.25)+
    slope.theme+
    #geom_text_repel(aes(color=factor(cons.area), label=name))+
    scale_color_manual(name = "Conservation area",
                       values=c("#7570b3", "#d95f02", "#1b9e77"),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    scale_shape_manual(name = "Conservation area",
                       values=c(15, 16, 17),
                       labels = c("Fir Butte", "Fern Ridge (FR)", "Willow Creek (WC)"))+
    annotate("text", x = 0.5, y = 0.4, label = "r = 0.91")+
    annotate("rect", xmin =-1.25, xmax = 2.25, ymin = -0.25, ymax = 0, alpha = 0.1)
  # + geom_text_repel(aes(label=name), color = "black")
change.ofs
```

### Put graphs together for FIGURE 4
A bit hard to see all panels in the knitted document, refer to graphs above. "r" aslo changed to "R^2" in Illustrator.
```{r, eval=FALSE}
(change.mu + change.fp + change.ofs) + guide_area() +
    plot_layout(guides="collect", ncol = 2) +
    plot_annotation(tag_levels = 'A') &
    theme(plot.tag = element_text(size = 14))
```


